
# {{ title }}


```{r}
pdat = remove_group_data(
  dat[!is.na(dat[["{{ iv1 }}"]]) & !is.na(dat[["{{ iv2 }}"]]),],
  "{{ dv.dif }}", c("{{ iv1 }}","{{ iv2 }}"))

pdat.long <- rbind(pdat[,c("id","{{ iv1 }}","{{ iv2 }}")],
                   pdat[,c("id","{{ iv1 }}","{{ iv2 }}")])
pdat.long[["{{ pivot.key }}"]] <- c(rep("pre", nrow(pdat)), rep("pos", nrow(pdat)))
pdat.long[["{{ pivot.key }}"]] <- factor(pdat.long[["{{ pivot.key }}"]], c("pre","pos"))
pdat.long[["{{ pivot.value }}"]] <- c(pdat[["{{ dv.pre }}"]], pdat[["{{ dv.pos }}"]])

y.position.min <- abs(
  max(pdat.long[["{{ pivot.value }}"]])
  - min(pdat.long[["{{ pivot.value }}"]]))/15

lvars = as.list(c("{{ dv.dif }}","{{ dv.pos }}","{{ dv.pre }}"))
names(lvars) = unlist(lvars)
```

## Pre-test and Post-test PairWise comparisons for: *{{ pivot.value }} ~ {{ iv1 }}:{{ iv2 }}*

```{r}
pwc.long <- group_by(pdat.long, {{ iv1 }}:{{ iv2 }}) %>%
  pairwise_wilcox_test({{ pivot.value }} ~ {{ pivot.key }}, detailed = T)

df <- pwc.long[,c(".y.","{{ iv1 }}:{{ iv2 }}","group1","group2","n1","n2","estimate",
                  "statistic","p.adj","p.adj.signif")]
```

```{r, echo=FALSE, purl=FALSE}
kable(df)
```

### Plot to compare pre- and post-test

```{r, dpi=300, fig.width={{ pfig.width }}, fig.height={{ pfig.height }}}
pwc.long <- group_by(pdat.long, {{ iv1 }}, {{ iv2 }}) %>%
  pairwise_wilcox_test({{ pivot.value }} ~ {{ pivot.key }}, detailed = T)

stat.test <- pwc.long %>% add_xy_position(x = "{{ pivot.key }}", fun = "mean_ci")
sidx = which(stat.test$p.adj.signif != "ns")
stat.test$y.position[sidx] <- stat.test$y.position[sidx] + y.position.min * (1:length(sidx))

gg <- ggline(
  pdat.long, x = "{{ pivot.key }}", y = "{{ pivot.value }}",
  color = "{{ iv2 }}", linetype = "{{ iv2 }}", shape = "{{ iv2 }}", size = 1.5,
  facet.by = "{{ iv1 }}", add = c("mean_ci"),
  position = position_dodge(width = 0.3), palette = color[["{{ iv2 }}"]])

pdat.long$xj = jitter(as.numeric(pdat.long[["{{ pivot.key }}"]]), amount=.1)
pdat.long$yj = jitter(pdat.long[["{{ pivot.value }}"]], amount = .01)

gg + geom_point(
  data = pdat.long, aes_string(x="xj",y="yj",colour="{{ iv2 }}"), size=0.5) +
  stat_pvalue_manual(
    stat.test, tip.length = 0, hide.ns = T, label.size = 5,
    position = position_dodge(width = 0.3), color = "{{ iv2 }}",
    label = "{ p.adj } ({ p.adj.signif })") + xlab("") + ylab("") +
  ylim(min(pdat.long$yj), max(pdat.long$yj)) +
  theme(strip.text = element_text(size = 14),
        axis.text = element_text(size = 14))
```

### New plot including percentagens in the differences


```{r, dpi=300, fig.width={{ pfig.width }}, fig.height={{ pfig.height }}}
stat.test <- pwc.long %>% add_xy_position(x = "{{ pivot.key }}", fun = "mean_ci")

stat.test$r <- sapply(abs(stat.test$estimate)/{{ max.y }}, FUN = function(x) {
   ifelse(x < 0.0001, "<1%", paste0(round(x*100,2), "%"))
})

gg <- ggline(
  pdat.long{{ plot.filter }}, x = "{{ pivot.key }}", y = "{{ pivot.value }}",
  color = "{{ iv2 }}", linetype = "{{ iv2 }}", shape = "{{ iv2 }}", size = 2,
  facet.by = "{{ iv1 }}", add = c("mean_ci"),
  position = position_dodge(width = 0.3), palette = color[["{{ iv2 }}"]]) +
  stat_pvalue_manual(
    stat.test, tip.length = 0, hide.ns = T, label.size = 5,
    color = "{{ iv2 }}",
    label = "{ r } ({ p.adj.signif })") + xlab("") + ylab("")

gg + theme(strip.text = element_text(size = 14),
           axis.text = element_text(size = 14))
```


## Scheirer and Wilcoxon PairWise comparisons for: *{{ pivot.value }} ~ {{ iv1 }}:{{ iv2 }}*


```{r}
sch <- lapply(lvars, FUN = function(x) {
  scheirer.test(pdat, x, c("{{ iv1 }}","{{ iv2 }}"), as.table = T) 
})
df <- do.call(rbind.fill, sch)
```

```{r, echo=FALSE, purl=FALSE}
kable(df)
```


```{r}
pwc <- lapply(lvars, FUN = function(x) {
  list(
    {{ iv1 }} = tryCatch(pairwise_wilcox_test(group_by(pdat, {{ iv2 }}),
                                 as.formula(paste0(x," ~ {{ iv1 }}")), detailed = T)
                         , error = function(e) NULL),
    {{ iv2 }} = tryCatch(pairwise_wilcox_test(group_by(pdat, {{ iv1 }}),
                                 as.formula(paste0(x," ~ {{ iv2 }}")), detailed = T)
                         , error = function(e) NULL)
  )
})

df <- do.call(rbind.fill, lapply(pwc, FUN =  function(x) {
  do.call(rbind.fill, x)
}))

ivs = c()
if ("{{ iv1 }}" %in% colnames(df)) ivs = c(ivs, "{{ iv1 }}")
if ("{{ iv2 }}" %in% colnames(df)) ivs = c(ivs, "{{ iv2 }}")
df <- df[,c(".y.",ivs,"group1","group2","n1","n2","estimate",
            "statistic","p.adj","p.adj.signif")]
```


```{r, echo=FALSE, purl=FALSE}
kable(df)
```

### Plot to compare results from pre and post

```{r}
plots <- lapply(lvars, FUN = function(y) {
  livs = list("{{ iv1 }}", "{{ iv2 }}")
  names(livs) = unlist(livs)
  lapply(livs, FUN = function(x) {
    iv2 = setdiff(names(livs), x)
    if (!is.null(pwc[[y]][[iv2]])) {
      stat.test <- pwc[[y]][[iv2]] %>% add_xy_position(x = x, fun = "max")
      sidx = which(stat.test$p.adj.signif != "ns")
      stat.test$y.position[sidx] <- stat.test$y.position[sidx] + y.position.min * (1:length(sidx))
      
      ggboxplot(pdat, x = x, y = y, fill = iv2, palette = color[[iv2]]) +
        stat_pvalue_manual(stat.test, tip.length = 0, hide.ns = T, label.size = 5,
                           label="{ p.adj } ({ p.adj.signif })") + xlab("")
    }
  })
})
```

```{r, dpi=300, fig.width={{ fig.width*2 }}, fig.height={{ fig.height }}}
if (!is.null(plots[["{{ dv.pre }}"]][["{{ iv1 }}"]]) &&
    !is.null(plots[["{{ dv.pos }}"]][["{{ iv1 }}"]])) {
  egg::ggarrange(plots[["{{ dv.pre }}"]][["{{ iv1 }}"]],
                 plots[["{{ dv.pos }}"]][["{{ iv1 }}"]], nrow = 1)  
}
```


```{r, dpi=300, fig.width={{ fig.width*2 }}, fig.height={{ fig.height }}}
if (!is.null(plots[["{{ dv.pre }}"]][["{{ iv2 }}"]]) &&
    !is.null(plots[["{{ dv.pos }}"]][["{{ iv2 }}"]])) {
  egg::ggarrange(plots[["{{ dv.pre }}"]][["{{ iv2 }}"]],
                 plots[["{{ dv.pos }}"]][["{{ iv2 }}"]], nrow = 1)
}
```

### Plot to compare differences (1st)

```{r, dpi=300, fig.width={{ fig.width }}, fig.height={{ fig.height }}}
psch = sch[["{{ dv.dif }}"]]
idx = which(psch$Effect == "{{ iv1 }}:{{ iv2 }}") 

dof = floor(as.double(psch$Df[idx]))
dof.res = floor(as.double(psch$Df[which(psch$Effect == "Residuals")]))
statistic = round(as.double(psch$H[idx]), 3)
p = round(as.double(psch[["p.value"]][idx]), 3)
pval = ifelse(p < 0.001,paste0(" , p<0.001"),paste0(" , p=",p))

if (!is.null(plots[["{{ dv.dif }}"]][["{{ iv1 }}"]]))
  plots[["{{ dv.dif }}"]][["{{ iv1 }}"]] +
    labs(subtitle = paste0("Scheirer-Ray-Hare H(", dof, ",", 
          dof.res, ")=", statistic, pval),
         caption = get_pwc_label(pwc[["{{ dv.dif }}"]][["{{ iv2 }}"]])) +
    ylab("{{ pivot.value }} (dif)") +
  theme(strip.text = element_text(size = 14),
        axis.text = element_text(size = 14))
```


### Plot to compare differences (1st)


```{r, dpi=300, fig.width={{ fig.width }}, fig.height={{ fig.height }}}
psch = sch[["{{ dv.dif }}"]]
idx = which(psch$Effect == "{{ iv1 }}:{{ iv2 }}") 

dof = floor(as.double(psch$Df[idx]))
dof.res = floor(as.double(psch$Df[which(psch$Effect == "Residuals")]))
statistic = round(as.double(psch$H[idx]), 3)
p = round(as.double(psch[["p.value"]][idx]), 3)
pval = ifelse(p < 0.001,paste0(" , p<0.001"),paste0(" , p=",p))

if (!is.null(plots[["{{ dv.dif }}"]][["{{ iv2 }}"]]))
  plots[["{{ dv.dif }}"]][["{{ iv2 }}"]] +
    labs(subtitle = paste0("Scheirer-Ray-Hare H(", dof, ",", 
          dof.res, ")=", statistic, pval),
         caption = get_pwc_label(pwc[["{{ dv.dif }}"]][["{{ iv1 }}"]])) +
    ylab("{{ pivot.value }} (dif)") +
  theme(strip.text = element_text(size = 14),
        axis.text = element_text(size = 14))
```


### Plot to compare differences using in one comparison

```{r, dpi=300, fig.width={{ pfig.width }}, fig.height={{ pfig.height }}}
psch = sch[["{{ dv.dif }}"]]
idx = which(psch$Effect == "{{ iv1 }}:{{ iv2 }}") 

dof = floor(as.double(psch$Df[idx]))
dof.res = floor(as.double(psch$Df[which(psch$Effect == "Residuals")]))
statistic = round(as.double(psch$H[idx]), 3)
p = round(as.double(psch[["p.value"]][idx]), 3)
pval = ifelse(p < 0.001,paste0(" , p<0.001"),paste0(" , p=",p))

dodge = 0.08
x.seg = sum(!is.na(unique(pdat[["{{ iv2 }}"]])))-1
d.seg = -1*dodge*x.seg/2


pwc1 = pwc[["{{ dv.dif }}"]][["{{ iv1 }}"]]
pwc2 = pwc1[pwc[["{{ dv.dif }}"]][["{{ iv1 }}"]]$p.adj.signif != "ns",]

if (!is.null(pwc1) && nrow(pwc2) > 0) {
  pwc2 = rstatix::add_xy_position(pwc2, dodge = dodge, fun = "mean_ci")
  
  for (f in sort(unique(pdat[["{{ iv2 }}"]]))) {
    fbool <- pwc2[["{{ iv2 }}"]] == f
    if (sum(fbool) > 0) {
      pwc2$xmin[which(fbool)] <- pwc2$xmin[which(fbool)]+d.seg
      pwc2$xmax[which(fbool)] <- pwc2$xmax[which(fbool)]+d.seg
    }
    d.seg <- d.seg + dodge
  }
} 

pwc1g <- pwc[["{{ dv.dif }}"]][["{{ iv2 }}"]]
pwc2g <- pwc1g[pwc[["{{ dv.dif }}"]][["{{ iv2 }}"]]$p.adj.signif != "ns",]

if (!is.null(pwc1g) && nrow(pwc2g) > 0) {
  pwc2g$y.position <- sapply(seq(1,nrow(pwc2g)), FUN = function(i) {
    rw <- as.list(pwc2g[i,])
    resp = -Inf
    for (atr2 in unique(pdat[["{{ iv2 }}"]])) {
      idx = which(pdat[["{{ iv1 }}"]] == rw[["{{ iv1 }}"]] & pdat[["{{ iv2 }}"]] == atr2)
      rmax = max(mean_ci(pdat[["{{ dv.dif }}"]][c(idx)]))
      if (rmax > resp) resp <- rmax
    }
    return(resp)
  })
  pwc2g$xpos <- sapply(seq(1,nrow(pwc2g)), FUN = function(i) {
    rw <- as.list(pwc2g[i,])
    tmp <- add_x_position(pwc[["{{ dv.dif }}"]][["{{ iv1 }}"]])
    min(tmp$xmin[which(tmp$group1 == rw[["{{ iv1 }}"]])],
        tmp$xmax[which(tmp$group2 == rw[["{{ iv1 }}"]])])
  })
  pwc2g$xmin <- pwc2g$xpos - abs(dodge*x.seg/2) 
  pwc2g$xmax <- pwc2g$xpos + abs(dodge*x.seg/2)
}

if (!is.null(pwc1) && nrow(pwc2) > 0) {
  pwc2$r <- sapply(abs(pwc2$estimate)/1, FUN = function(x) {
     ifelse(x < 0.0001, "<1%", paste0(round(x*100,2), "%"))
  })
}

pd <- ggplot2::position_dodge(width = sum(!is.na(unique(pwc2[["{{ iv2 }}"]])))*dodge)

lp <- ggpubr::ggline(pdat, x="{{ iv1 }}", y = "{{ dv.dif }}", color = "{{ iv2 }}", linetype = "{{ iv2 }}",
                       palette = color[["{{ iv2 }}"]], plot_type='b', size=2,
                       position = pd, add = "mean_ci", ylab = "")

if (!is.null(pwc1) && nrow(pwc2) > 0)
  lp <- lp + ggpubr::stat_pvalue_manual(pwc2, color = "{{ iv2 }}", linetype = "{{ iv2 }}",
                                          hide.ns = T, tip.length = 0,
                                          label = "{ r } ({ p.adj.signif })")

if (!is.null(pwc1g) && nrow(pwc2g) > 0) {
  y.pos.min = max(pwc2g$y.position)/10
  if (nrow(pwc2) > 0)
    y.pos.min = max(pwc2$y.position, pwc2g$y.position)/10
  
  pwc2g$r <- sapply(abs(pwc2g$estimate)/1, FUN = function(x) {
     ifelse(x < 0.0001, "<1%", paste0(round(x*100,2), "%"))
  })
  pwc2g$y.position <- pwc2g$y.position + y.pos.min
  
  for (i in which(pwc2g$p.adj < 0.05)) {
    x1 = pwc2g$xmin[i]
    x2 = pwc2g$xmax[i]
    y.pos = pwc2g$y.position[i]
    label = pwc2g$p.adj.signif[i]
    label = paste0(pwc2g$r[i]," (",label,")")
    
    lp <- lp + ggplot2::geom_segment(x = x1, y = y.pos, xend = x2, yend = y.pos) +
      ggplot2::geom_text(x=(x1+x2)/2, y = y.pos+y.pos.min/3, label=label)
  }
}

lp + labs(subtitle = paste0("Scheirer-Ray-Hare H(", dof, ",", 
                            dof.res, ")=", statistic, pval),
          caption = get_pwc_label(pwc[["{{ dv.dif }}"]][["{{ iv2 }}"]])) +
  xlab("") + ylab("") +
  theme(strip.text = element_text(size = 14),
        axis.text = element_text(size = 14))
```


